---
title: "My code repository"
author: "Ishan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic codes

## Group|mean|relocate
```{r}
data_mean <- data %>% 
  group_by(Cycle,Mycode,Code,District,FullDate,Name) %>% 
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = T))) %>% 
  relocate(Cycle,Mycode,Code,District,FullDate,Name, .after = last_col())


setdiff(colnames(data_mean),colnames(data.PCA))
```
## mutate
```{r}
data.sup <- data.sup %>% mutate(Eel = Eel/(Surf.area*100))
```

## Subset
```{r}
subset(data, Phase == "During")
```


## Custom labels
```{r}
#auto arrangement
cat(paste0(colnames(data), ' = " "', collapse = " , "))

custom_labels <- c(
  Temperature = "Temp",
  DO = "DO",
  EC = "EC",
  TDS = "TDS",
  Water_pH = "Water-pH",
  Water_ORP = "Water-ORP")
```

## Arrange colnames
```{r}
#for selection
cat(paste0("(" ,colnames(data), ")", collapse = " , "))

#for selection
cat(paste0('"' ,colnames(data), '"', collapse = " , "))

#for GAM
cat(paste0("s(" ,summary_table1$Variable, ")", collapse = " + "))
```

# Distribution check | Histograms

```{r}
names1 <- colnames(data)

for (nam in names1) {
  if (is.numeric(data[[nam]])) {
    hist(data[[nam]], main = nam, breaks = 30)
  }
}
```

# Normality check
```{r}
# Shapiro-Wilk Normality Test for all numeric columns with conclusion
normality_summary <- data.frame(
  Column = character(),
  P_Value = numeric(),
  Conclusion = character(),
  stringsAsFactors = FALSE
)

for (col_name in names(data.trf)) {
  column_data <- data[[col_name]]
  
  if (is.numeric(column_data) && length(column_data) >= 3 && length(column_data) <= 5000) {
    test_result <- shapiro.test(column_data)
    p_val <- test_result$p.value
    conclusion <- ifelse(p_val > 0.05, "Normal", "Not Normal")
    
    normality_summary <- rbind(normality_summary, 
                               data.frame(Column = col_name, 
                                          P_Value = round(p_val, 5), 
                                          Conclusion = conclusion))
  }
}

# View summary table
print(normality_summary)


```


# Correlation

## Test
```{r}
t1 <- cor.test(data$Builtup.area, data$Agri.area)
t1$p.value     # p-value 
t1$estimate    # r
t1$estimate^2  # R2
t1
```


## Correlation matrix Plot | Circles
```{r  fig.width= 8, fig.height=7, dpi= 300}
library(corrplot)

# First correlation plot for Percentage LULC
cols_of_interest <- c("Forest", "Agriculture", "Builtup", "Water")
data_subset <- data[cols_of_interest]
cor_matrix <- cor(data_subset, use = "complete.obs") 

# Plotting the correlation matrix for Percentage LULC
corrplot::corrplot(
  cor_matrix, 
  method = "circle",
  type = "lower",
  tl.col = "black",
  tl.srt = 45,
  tl.cex = 1,
  addCoef.col = "black",
  number.cex = 0.9,
  title = "Percentage LULC", 
  mar = c(0, 0, 2, 0),
  col = colorRampPalette(c("blue", "white", "red"))(100),
  diag = FALSE,
  cl.ratio = 0.2,          # Width of color legend
  cl.align.text = "l",     # Align color legend text
  cl.cex = 0.8             # Color legend text size
)
```

## Correlation matrix Plot | Lower Traingle
```{r  fig.width= 8, fig.height=7}

# Calculate R² matrix
cor_matrix_r2 <- cor_matrix_original^2
colnames(cor_matrix_r2) <- custom_labels
rownames(cor_matrix_r2) <- custom_labels

# Create a copy and set LOWER triangle to NA to only show UPPER half
cor_matrix_r2_upper <- cor_matrix_r2
cor_matrix_r2_upper[lower.tri(cor_matrix_r2_upper, diag = TRUE)] <- NA

# Melt the matrix
cor_melt <- melt(cor_matrix_r2_upper, na.rm = TRUE, varnames = c("Var1", "Var2"), value.name = "R2")

# Set factor levels to maintain order
cor_melt$Var1 <- factor(cor_melt$Var1, levels = custom_labels)
cor_melt$Var2 <- factor(cor_melt$Var2, levels = rev(custom_labels))  # Reverse for proper orientation

# Set font for Windows
if (.Platform$OS.type == "windows") {
  windowsFonts(Times = windowsFont("Times New Roman"))
}

# Create the plot with only UPPER half
ggplot(cor_melt, aes(x = Var1, y = Var2, fill = R2)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", R2)), size = 4, fontface = "bold", family = "Times") +
  scale_fill_gradient2(
    low = "white", mid = "salmon", high = "red", midpoint = 0.5,
    limits = c(0, 1), na.value = "white"
  ) +
  labs(fill = expression(R^2)) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = "bold", family = "Times"),
    axis.text.y = element_text(face = "bold", family = "Times"),
    legend.title = element_text(face = "bold", family = "Times"),
    legend.text = element_text(family = "Times"),
    plot.title = element_text(face = "bold", family = "Times"),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  ) +
  ggtitle("R² Correlation Matrix (Upper Triangle)") +
  scale_x_discrete(limits = custom_labels) +
  scale_y_discrete(limits = rev(custom_labels))
```

## Correlation network
```{r fig.width= 6, fig.height=3, dpi= 300}
library(igraph)
library(ggraph)
library(dplyr)

var.selected <- c("Forest", "Agriculture", "Builtup", "Water")

# Select LULC columns
selected_data <- dplyr::select(data, dplyr::all_of(var.selected))

# Compute correlation
cor.tab <- cor(selected_data[sapply(selected_data, is.numeric)], use = "pairwise.complete.obs")

# Convert correlation matrix to table and filter
filtered_cor <- as.data.frame(as.table(cor.tab)) %>%
  dplyr::rename(Var1 = Var1, Var2 = Var2, Correlation = Freq) %>%
  dplyr::filter(abs(Correlation) >= 0.7, Var1 != Var2) %>%
  dplyr::arrange(desc(Correlation))

# Create edges
cor_edges <- dplyr::select(filtered_cor, Var1, Var2, Correlation)

# Graph object
cor_graph <- graph_from_data_frame(cor_edges, directed = FALSE)

# Add labels as node attribute
V(cor_graph)$label <- custom_labels[V(cor_graph)$name]

# Plot
ggraph(cor_graph, layout = "kk") +
  geom_edge_link(aes(edge_alpha = abs(Correlation), edge_width = abs(Correlation), edge_color = Correlation)) +
  scale_edge_color_gradient2(low = "salmon3", mid = "white", high = "gold3", midpoint = 0, name = "Correlation") +
  scale_edge_width_continuous(name = "Correlation") +
  scale_edge_alpha_continuous(name = "Correlation") +
  geom_node_point(size = 2, color = "black") +
  geom_node_text(aes(label = label), repel = TRUE, size = 3) +
  theme_void() +
  theme(legend.position = "right", 
        legend.box = "vertical", 
        plot.margin = ggplot2::margin(10, 10, 10, 10),
        plot.title = element_text(hjust = 0.5, vjust = 4, size = 13, face = "bold"))

```

## Hierarchiacal clustering of variables based on correlation

```{r}
library(ggdendro)
library(dendextend)

# Compute correlation matrix and convert to distance matrix
cor_matrix <- cor(data[, env.var], use = "complete.obs")
dist_matrix <- as.dist(1 - cor_matrix)  # Use 1 - correlation directly for distance

# Perform hierarchical clustering
hclust_result <- hclust(dist_matrix, method = "ward.D2")

# Convert the result to a dendrogram object
dend <- as.dendrogram(hclust_result)

# Replace labels with custom labels
labels(dend) <- custom_labels[labels(dend)]

# Highlight branches based on correlation threshold
cor_threshold <- 0.6
cor_pairs <- which(abs(cor_matrix) > cor_threshold, arr.ind = TRUE)

highlight_branches <- function(dend, cor_pairs) {
  for (i in 1:nrow(cor_pairs)) {
    idx1 <- cor_pairs[i, 1]
    idx2 <- cor_pairs[i, 2]
    dend <- dendextend::set(dend, "branches_col", value = "red", k = 7)
  }
  return(dend)
}

# Apply highlighting
dend <- highlight_branches(dend, cor_pairs)

# Customize and plot the dendrogram
dend <- dend %>%
  color_branches(k = 7) %>%  # Color branches for 7 clusters
  set("labels_cex", 0.8)     # Adjust label size

par(mar = c(8, 3, 4, 4))  # Adjust margins for better label visibility

# Adjust y-axis to display correlation
plot(dend, 
     main = "    Hierarchical Clustering of Variables", 
     ylab = "1 - Correlation", 
     ylim = c(0, 0.5))
abline(h = 0.4, col = "gray", lty = 2, lwd = 2)

# Add a secondary y-axis with correlation values
axis(side = 4, at = seq(0, 1, by = 0.2), 
     labels = round(1 - seq(0, 1, by = 0.2), 2))
mtext("Correlation", side = 4, line = 3)

```

# Variance

## Variance analysis
```{r fig.width=7, fig.height=3}
library(corrr)
library(tidymodels)

library(showtext)
showtext_auto()
font_add("Times", "C:/Windows/Fonts/times.ttf")  # adjust if needed

# PCA preparation recipe
pca_prep_recipe <- recipe( ~ ., data = data[, env.var]) %>%
  step_nzv(all_numeric_predictors()) %>%
  prep()

# Apply the recipe to create the PCA-ready data1set
pca_ready_df <- bake(pca_prep_recipe, new_data = NULL)

# Compute variance and arrange
na_filter <- pca_ready_df %>% 
  summarize(across(everything(), ~ var(scale(., center = FALSE), na.rm = TRUE))) %>% 
  pivot_longer(cols = everything(), names_to = "feature", values_to = "variance") %>%  
  arrange(desc(variance)) %>%
  mutate(variance = round(variance, 2))  # Round variance here

# Add readable labels from custom_labels
na_filter <- na_filter %>%
  mutate(label = custom_labels[feature])

# Plot with new labels
p <- ggplot(na_filter, aes(x = reorder(label, -variance), y = variance)) +
  geom_point(shape = 16, color = "black", size = 2) +  # fixed (no stat="identity")
  labs(
    x = "Parameter",
    y = "Variance"
  ) +
  theme_bw(base_family = "Times") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    panel.grid = element_blank(),
  )
p
# View variance table with readable labels
print(na_filter)

# Print labels for use elsewhere
cat(paste0('"', na_filter$label, '"', sep = " ,"))

ggsave(
  filename = file.path("E:/1. Postgraduate (MPhill)/01. Thesis/3. Publications/1. GFP/2. Manuscipts/Other figures", "FIG02.pdf"),
  plot = p,
  device = cairo_pdf
)


```

# Clustering

## Basic Dendogram
```{r}
# Obtain Distance matrix
env.dist <- dist(scale(data), method = "euclidean")

# get clusters based on distance
env.clust <- hclust(env.dist, method = "ward.D2")

# limit number of clusters for L bow plot
max_clusters <- 15
wss <- numeric(max_clusters)

# Looping data for L bow plot
for (k in 1:max_clusters) {
  cluster_assignments <- cutree(env.clust, k)
  wss[k] <- sum(sapply(unique(cluster_assignments), function(clust) {
    sum(dist(scale(data.PCA)[cluster_assignments == clust, ])^2)
  }))
}

# Plotting L Bow plot
plot(1:max_clusters, wss, type = "b", pch = 19,
     xlab = "Number of Clusters",
     ylab = "Within-Cluster Sum of Squares (WSS)",
     main = "Elbow Method for Optimal Clusters")

# Plotting the dendogram
plot(env.clust, labels = F, xlab = " ")

# mark clusters based on L Bow plot
rect.hclust(env.clust, k = 4, border = c("#ff7f0e","black", 
                                         "purple","green4","pink2", "gold3"))
```

## MDS clustered by dendogram
```{r}

# Cut tree at any point
env.cl2 <- cutree(env.clust,2)
env.cl3 <- cutree(env.clust,3)
env.cl4 <- cutree(env.clust,4)
env.cl5 <- cutree(env.clust,5)
env.cl6 <- cutree(env.clust,6)

env.mds <- cmdscale(env.dist)

env.mds <- as.data.frame(env.mds)

# Plot MDS
plot(env.mds[,1], env.mds[,2], type = "n", xlab = "MDS1", ylab = "MDS2")
points(env.mds, pch = 21, bg = c("#ff7f0e","black", 
                                 "purple","green4","pink2","gold3")[env.cl5], cex = 1.5)
```


# Boxplot with Response parameter
```{r}
ggplot(data, aes(x = Code, y = GFP_stocking, fill = Code)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "GFP stocking across reservoirs",
       x = "Cluster", y = "GFP stocking") +
  theme_minimal() +
  theme(legend.position = "none")
```


# PCA

## Basic PCA
```{r}
pca_1 <- princomp(scale(data[,2:ncol(data)]))

# Access % of variance explained 
pv <- (pca_1$sdev^2 / sum(pca_1$sdev^2))*100   

# PCA Biplot visualisation 
biplot(pca_1, xlabs = data$Mycode, col = c("gold2","blue"),
       xlab= paste0("PC I (",round(pv[1],1),"% )"), xlim = c(-0.2,0.2),
       ylim = c(-0.2,0.14), cex = 0.8,
       ylab= paste0("PC II (",round(pv[2],1),"% )"))+
  abline(h = 0, v = 0, col = "gray30", lty = 8)

```

## PCA using RDA
```{r}
library(vegan)
library(ggplot2)

# Generate PCA plots using Vegan package ------------------
env.pca <- rda(scale(data))
data<- data %>%
  ungroup() %>%                        
  mutate(clas2 = env.cl2,
         clas3 = env.cl3,
         clas4 = env.cl4,
         clas5 = env.cl5,
         clas6 = env.cl6,
         PC_I = env.pca$CA$u[,1],
         PC_II = env.pca$CA$u[,2],
         PC_III = env.pca$CA$u[,3],
         PC_IV = env.pca$CA$u[,4],
         PC_V = env.pca$CA$u[,5]) %>%   
  group_by(Cycle, Mycode, Code, District, FullDate, Name, Season, 
           Inlet_control, Monsoon, Lithology, Basin)

#saveRDS(data.sup, "Data/2.Data_all.rds")

################ analysis -------------------------
x <- 1
y <- 2

# Obtain percentage variance explained automatically using pca table
var.exp <- round(env.pca$CA$eig / sum(env.pca$CA$eig) * 100,1)
names(var.exp) <- as.character(as.roman(1:length(var.exp)))

# Obtain percentage variance explained automatically using pca table
xlabel <- function(x) {paste0("PC ", names(var.exp[x]), " (", var.exp[x], "%)")}
ylabel <- function(y) {paste0("PC ", names(var.exp[y]), " (", var.exp[y], "%)")}

# Create data frames from the relevant parts of env.pca
df_u <- as.data.frame(env.pca$CA$u)
df_v <- as.data.frame(env.pca$CA$v)

# Ensure the data frames have appropriate column names for ggplot
colnames(df_u) <- paste0("PC", 1:ncol(df_u))
colnames(df_v) <- paste0("PC", 1:ncol(df_v))


# Plotting with ggplot
ggplot() +
  geom_point(data = df_u, aes(x = PC1, y = PC2), color = "black", size = 2) +
  #stat_ellipse(data = df_u, aes(x = PC1, y = PC2, color = as.factor(data.sup$Monsoon)), type = "t", level = 0.95, linewidth = 1) +
  geom_text_repel(data = df_v, aes(x = 1.1*PC1, y = 1.1*PC2, label = rownames(df_v)), size = 3.5, color = "blue3", fontface = "bold", box.padding = 0.3, point.padding = 0.3) +
  geom_segment(data = df_v, aes(x = 0, y = 0, xend = 1.1*PC1, yend = 1.1*PC2), arrow = arrow(length = unit(0.2, "cm")), color = "lightblue4") +
  geom_vline(xintercept = 0, color = "grey70", linetype = 2) +
  geom_hline(yintercept = 0, color = "grey70", linetype = 2) +
  scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
  scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
  xlab(xlabel(1)) +
  ylab(ylabel(2)) +
  labs(fill = "Total Production (kg/ha)") +
  scale_color_manual(values = c("#2CCACE", "#C77CFF","#F8766D","#83B20D")) +
  labs(color = "Monsoon Period") +
  theme(
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(color = "gray25"),
    text = element_text(size = 12),
    axis.text = element_text(size = 10, colour = "gray25"),
    axis.title = element_text(size = 14, colour = "gray25"),
    legend.title = element_text(size = 12),
    legend.key = element_blank(),
    legend.position = "right",
    legend.text = element_text(size = 12, color = "black")
  )
```

## PCA with Ordiplot
```{r}
# Convert data to data frame and set row names
data.PCA <- as.data.frame(data.PCA)
rownames(data.PCA) <- data$Mycode

# Perform PCA
pca.test <- rda(scale(data.PCA))

# Extract plot data
plot1 <- ordiplot(pca.test, choices = c(1, 2))
sites.long1 <- sites.long(plot1, env.data = data.PCA)
species.long1 <- species.long(plot1)
axis.long2 <- axis.long(pca.test, choices = c(1, 2))

# Generate ordisurf data
A1.surface <- ordisurf(plot1, y = data$Total_Hrvst, nlevels = 9)
A1.grid <- ordisurfgrid.long(A1.surface)

# Filter finite values for contour
finite_z <- A1.grid %>% filter(is.finite(z))
max_z <- max(finite_z$z)
contour_breaks <- seq(0, max_z + 50, by = 50)  # Adjust intervals as needed
A1.grid <- A1.grid %>% filter(is.finite(z))

# Ensure Monsoon is a factor
#data.sup$Monsoon <- as.factor(data.sup$Monsoon)
data.sup$Monsoon <- factor(data.sup$Monsoon, levels = c("Northeast", "1st_Inter_Monsoon", "Southwest","2nd_Inter_Monsoon"))

####### PCA plot with Production surface  -----------
ggplot() +
  geom_contour_filled(data = A1.grid, aes(x = x,y = y,z = z),
                      breaks = seq(0, max_z + 50, by = 50)) +
  scale_fill_brewer(palette = "Blues") +
  stat_ellipse(data = sites.long1, 
               aes(x = as.numeric(axis1), y = as.numeric(axis2), 
                   color = as.factor(data.sup$Monsoon)), 
               type = "t", level = 0.95, linewidth = 1) +
  # geom_text_repel(data = species.long1, 
  #                 aes(x = 1.1*as.numeric(axis1), y = 1.1*as.numeric(axis2), 
  #                     label = labels), 
  #                 size = 3.5, color = "red3", fontface = "bold", 
  #                 box.padding = 0.3, point.padding = 0.3) +
  geom_vline(xintercept = 0, color = "grey70", linetype = 2) +
  geom_hline(yintercept = 0, color = "grey70", linetype = 2) +
  scale_x_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
  scale_y_continuous(sec.axis = dup_axis(labels = NULL, name = NULL)) +
  xlab(axis.long2[1, "label"]) +
  ylab(axis.long2[2, "label"]) +  
  labs(fill = "Total Production (kg/ha)") +
  scale_color_manual(values = c("#2c51c9", "#C78CFF","#fa9750","#6b8f10")) +
  labs(color = "Monsoon Period", fill = "Total Production (kg/ha)") +
  theme(
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(color = "gray25"),
    text = element_text(size = 12),
    axis.text = element_text(size = 10, colour = "gray25"),
    axis.title = element_text(size = 14, colour = "gray25"),
    legend.title = element_text(size = 12),
    legend.key = element_blank(),
    legend.position = "right",
    legend.text = element_text(size = 12, color = "black")
  )

```

# Cluster statistics

## permutational dispersions
```{r}


group <- as.factor(data$hclust_cluster)

library(vegan)

# Standardize environmental data
env.dist <- vegdist(scale(data.pca), method = "euclidean")  # or "bray" if appropriate

# Run PERMANOVA
permanova_result <- adonis2(env.dist ~ group, permutations = 999)
print(permanova_result)


# Test dispersion
dispersion <- betadisper(env.dist, group)

# ANOVA on the dispersions
permdisp_result <- anova(dispersion)
print(permdisp_result)

# Optional: pairwise tests
permdisp_pairwise <- permutest(dispersion, permutations = 999)
print(permdisp_pairwise)

# Optional: visualize
plot(dispersion)


```

## ANOSIM
```{r}
library(vegan)

# Prepare the distance matrix (use scaled data)
env.dist <- vegdist(scale(data.pca), method = "euclidean")  # or method = "bray"

# Ensure grouping factor
group <- as.factor(data$kmeans5)

# Run ANOSIM
anosim_result <- anosim(env.dist, group, permutations = 999)
print(anosim_result)

# Optional plot
plot(anosim_result)

```

## SIMPER
```{r}
library(vegan)

# Environmental data (already scaled is best)
env.scaled <- scale(data.pca)  # assuming data.pca = data.mean[, var.env]

# Grouping factor
group <- as.factor(data$kmeans5)

# Run SIMPER
simper_result <- simper(env.scaled, group)

# View summary
summary(simper_result)

# Example: get top contributors between groups 1 and 2
simper_result[[1]]  # or simper_result[["1_2"]] depending on how it's named

```

# GAM

## basic GAM
```{r}
gam1 <- gam(EC ~ s(Agri.std), data = data,, method = "REML")

summary(gam1)

plot(gam1)

gam.check(gam1)

```

## GAM plot ggplot
```{r}
ggplot(data, aes(x = Agri.std, y = EC, color = Phase, fill = Phase)) +
    geom_point(size = 1.2, alpha = 0.3) +
    geom_smooth(
      method = "gam",
      method.args = list(method = "REML"),
      formula = y ~ s(x),
      se = TRUE,
      linewidth = 1.2, 
      alpha = 0.09
    ) +
    scale_color_manual(values = phase_colors) +
    scale_fill_manual(values = phase_colors) +
    labs(x = "Agri.std", y = custom_labels0[variable]) +
    theme_classic(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12, face = "bold"),
      axis.title = element_text(size = 14, face = "bold"),
      legend.title = element_text(size = 12, face = "bold"),
      legend.text = element_text(size = 11)
    )
  
```

## Check GAM variation for all parameters individually

```{r}
library(mgcv)
library(tibble)
library(ggplot2)
library(gratia)

# "Catchment.area", "Surf.area" excluded beacause I already use the suraface area to standardize the catch and catchment area also highly correlates with it.

env.var <- c("Temperature", "DO", "log(EC)", "Water_pH", "Water_ORP", "log(Chl_a)", "SDD", "P_Alk", "T_Alk", "Water_TP", "Water_TN", "Sed_TP", "log(Sed_TN)", 
             "Sed_Col_L", "Sed_Col_a",  "Sed_OM", "Sed_pH", "Sed_ORP", 
             "Max.depth", "Tank.Elevation")

summary_table <- tibble(
  Variable = character(),
  edf = numeric(),
  Ref_df = numeric(),
  F_value = numeric(),
  p_value = numeric(),
  R_squared = numeric(),
  AIC = numeric()
)

model_results <- list()

for (X in env.var) {
  formula_text <- paste("log(GFP_Hrvst + 0.1) ~ s(", X ,", k = 15)", sep = "")
  this_formula <- as.formula(formula_text)
  
  # Try-catch to handle any modeling issues
  try({
    gam_model <- gam(this_formula, data = data, method = "REML")
    model_results[[X]] <- gam_model
    
    s <- summary(gam_model)
    
    if (!is.null(s$s.table)) {
      edf <- round(s$s.table[1, "edf"],2)
      ref.df <- round(s$s.table[1, "Ref.df"],2)
      f_val <- round(s$s.table[1, "F"],3)
      p_val <- round(s$s.table[1, "p-value"], 3)
    } else {
      edf <- ref.df <- f_val <- p_val <- NA
    }
    
    r2 <- round(s$r.sq, 3)
    model_aic <- AIC(gam_model)
    
    summary_table <- add_row(summary_table,
                             Variable = X,
                             edf = edf,
                             Ref_df = ref.df,
                             F_value = f_val,
                             p_value = p_val,
                             R_squared = r2,
                             AIC = model_aic)
  }, silent = TRUE)
}

View(summary_table)

#write.xlsx(summary_table, "gam_summary_table.xlsx", row.names = FALSE)

```

## Univariate GAM loop
```{r}
env.var <- c("Temperature", "DO", "log(EC)", "Water_pH", "Water_ORP", "log(Chl_a)", "SDD", "P_Alk", "log(T_Alk)", "log(Water_TP)", "Water_TN", "Sed_TP", "log(Sed_TN)", "Sed_Col_L", "Sed_Col_a", "Sed_OM", "Sed_pH", "Sed_ORP", "log(Max.depth)", "Tank.Elevation", "log(Surf.area)", "GFP_stocking")

custom_labels1 <- c(
  Temperature = "Temperature",
  DO = "DO",
  `log(EC)` = "log(EC)",
  TDS = "TDS",
  Water_pH = "Water-pH",
  Water_ORP = "Water-ORP",
  `log(Chl_a)` = "log(Chl-a)",
  SDD = "SDD",
  P_Alk = "P-Alk",
  `log(T_Alk)` = "log(T-Alk)",
  Hardness = "Hardness",
  `log(Water_TP)` = "log(Water-TP)",
  Water_TN = "Water-TN",
  Sed_TP = "Sed-TP",
  `log(Sed_TN)` = "log(Sed_TN)",
  Sed_Col_L = "Sed-Col-L",
  Sed_Col_a = "Sed-Col-a",
  Sed_Col_b = "Sed-Col-b",
  Sed_OM = "Sed-OM",
  Sed_pH = "Sed-pH",
  Sed_ORP = "Sed-ORP",
  Surf.area = "Surf area",
  Perimeter = "Perimeter",
  `log(Max.depth)` = "log(max-depth)",
  Catchment.area = "Catchment area",
  Tank.Elevation = "Altitude",
  Atm.pressure = "Atm-pressure",
  GFP_stocking = "GFP-stocking",
  `log(Surf.area)` = "log(Surf.area)"
)

for (para in env.var) {
  formula <- as.formula(paste("log(GFP_Hrvst + 0.1) ~ s(", para, ")", sep = ""))
  gam1 <- mgcv::gam(formula, data = data, method = "REML", select = TRUE)
  
  print(paste("Results for", para))
  print(summary(gam1))
  #gam.check(gam1)
  
  plot(gam1, shade = TRUE, shade.col = "lightblue", rug = TRUE, 
       residuals = TRUE, cex = 2, 
       xlab = custom_labels1[para], 
       ylab = paste("s(", custom_labels1[para], ")", sep = ""))
}

```

# Random Forest

```{r  fig.width= 5, fig.height=3.6}
library(randomForest)
library(rpart)
library(dplyr)
set.seed(22)

env.var <- c("Temperature", "DO", "EC", "TDS", "Water_pH", "Water_ORP", "Chl_a", "SDD", "TSS", "P_Alk", "T_Alk", "Hardness", "Water_TP", "Water_TN", "Sed_TP", "Sed_TN", "Sed_Col_L", "Sed_Col_a", "Sed_Col_b", "Sed_OM", "Sed_pH", "Sed_ORP")

# Create a temporary copy and rename GFP_Hrvst to GFP_catch
temp_data <- data %>%
  ungroup() %>%
  select(LULC_pca1, all_of(env.var))
# Remove missing values
temp_data <- na.omit(temp_data)

# Run Random Forest model
rfmodel <- randomForest(LULC_pca1 ~ ., data = temp_data, importance = TRUE, ntree = 1001, mtry = 5)

# Extract importance matrix
importance_matrix <- randomForest::importance(rfmodel)

# Get top 10 by %IncMSE and IncNodePurity
top_10_mse <- tail(importance_matrix[order(importance_matrix[ , "%IncMSE"]), ], 10)
top_10_purity <- tail(importance_matrix[order(importance_matrix[ , "IncNodePurity"]), ], 10)

varImpPlot(rfmodel, main = "Water quality and seidment quality")

dotchart(top_10_mse[ , "%IncMSE"],
         labels = custom_labels0[rownames(top_10_mse)],
         xlab = "% Increase in MSE",
         main = "LULC_pca1",
         cex = 0.8)

# Print model summary
print(rfmodel)

```

# Linear regression

```{r}
data2 <- subset(data, Phase == "During")

gam1 <- lm(T_Alk ~ Agri.std, data = data)

summary(gam1)

car::avPlots(gam1)
```
































